MS_Intro_Short

### 2024-02-25 Markdown/Manuscript/Child_RMDs/MS_Imp_Example.Rmd
on to discussion

## Model user requirements
The initial model user requirements (MURs) that we specified for initial readyforwhatsnext modules we developed for utility mapping were:

@@

MUR 2) *Dataset description tools* that: (i) attach metadata to a human records dataset that specifies the unique person identifier, group assignment and data collection round variables; and (ii) generate tabular and graphical summaries of dataset descriptive statistics.

@@

MUR 4) *Utility mapping model construction tools* that: (i) attach structured metadata about the target utility instrument and candidate predictor variables and covariates to a human records dataset; (ii) generate descriptive tables, a correlation matrix and plots about the variables to be assessed for inclusion in models; (iii) attach structured metadata about the candidate utility mapping models to be assessed; (iv) generate tabular and graphical summaries of the performance of each combination of model, predictors and covariates specified for evaluation by either a user or by a default algorithm; (v) strip confidential records from the constructed models so that they are able to be publicly shared.

>>>>

## Model user requirements
The initial model user requirements (MURs) that we specified for the readyforwhatsnext modules we developed for utility mapping are:

@@

MUR 2) *Dataset description tools* that: (i) attach structured metadata to a human records dataset that specify the unique person identifier, group assignment and data collection round variables; and (ii) generate tabular and graphical summaries of dataset descriptive statistics.

@@

MUR 4) *Utility mapping model construction tools* that: (i) attach structured metadata about the target utility instrument and candidate predictor variables and covariates to be used in constructing utility mapping models to a human records dataset; (ii) generate descriptive tables, a correlation matrix and plots about the variables to be assessed for inclusion in models; (iii) attach structured metadata about the candidate utility mapping models to be assessed; (iv) generate tabular and graphical summaries of the performance of each combination of model, predictors and covariates specified for evaluation by either a user or by a default algorithm; (v) strip confidential records from the constructed models so that they are able to be publicly shared.

### 2024-02-25 Markdown/Manuscript/Child_RMDs/MS_Imp_Motivation.Rmd
on to discussion

We became interested in modular and open-source approaches to CHEM development in 2018 when beginning to plan how to synthesise a number of current or planned economic modelling projects in youth mental health in one multi-purpose model. This overarching youth mental health model is called readyforwhatsnext [@rfwn2024].

## Economic topics
Initially, we are constructing readyforwhatsnext from projects in four of the twelve domains of health economics identified by Wagstaff and Culyer [@wagstaff2012four]:

@@

Potential future directions could incorporate projects in two additional Wagstaff and Culyer domains of: (i) public health (to model the impact of selected fiscal policy and regulation options on young people's mental health); and (ii) human resources (to model the supply and behaviours of the youth mental health workforce). Our ultimate aim is for readyforwhatsnext to flexibly combine its constituent parts in analyses that help answer questions in two additional Wagstaff and Culyer domains:

@@

## Transferability
We are principally interested in using readyforwhatsnext to answer policy questions relating to the mental health of young people in Australia. Australian mental health services are primarily planned and commissioned by the Australian Government, the governments of Australia's state and territories and independent organisations focused on smaller regions called Primary Health Networks. Promoting mentally healthy local communities is also an important role for Australian local governments. Australia is an international leader in developing innovative youth mental health service models that are subsequently adopted in other jurisdictions [@RN27]. For these reasons, we want readyforwhatsnext to be implemented in a manner that facilitates its transfer to multiple decision contexts.

However, there are ethical risks associated with each potential transfer, in particular the need to ensure social acceptability, fitness for purpose and appropriate use of readyforwhatsnext in each new decision context. For that reason, we have set the goal of readyforwhatsnext meeting six explicit criteria for transparency, reusability and updatability (TRU criteria). We have describe these criteria and their origins in ethical modelling practice elsewhere [@ethicsCHEMS2024], but briefly they are:

@@

U6) Systematic retesting and deprecation.

>>>>

We became interested in solving some of the technical barriers to transferable CHEMs in 2018 when considering how to synthesise multiple youth mental health economic modelling projectsin one multi-purpose model. This overarching youth mental health model is called readyforwhatsnext [@rfwn2024].

## Economic topics
We are constructing readyforwhatsnext from projects in four of the twelve domains of health economics identified by Wagstaff and Culyer [@wagstaff2012four]. These are:

@@

Potential future directions for readyforwhatsnext may incorporate projects in two additional Wagstaff and Culyer domains of: (i) public health (to model the impact of selected fiscal policy and regulation options on young people's mental health); and (ii) human resources (to model the supply and behaviours of the youth mental health workforce). Our ultimate aim is for readyforwhatsnext to flexibly combine its constituent parts in analyses that help answer questions in two additional Wagstaff and Culyer domains:

@@

## Transferability
We are principally interested in using readyforwhatsnext to answer policy questions relating to the mental health of young people in Australia. Australian public mental health services can be planned and commissioned by the Australian Government, the governments of Australia's state and territories and independent organisations focused on smaller regions called Primary Health Networks. Promoting mentally healthy local communities is also an important role for Australian local governments. The mental health of young people is a global priority [@https://doi.org/10.1002/wps.20938] and Australia has initiated youth mental health service model reforms that have been adopted in other jurisdictions [@RN27]. For these reasons, we aim to develop readyforwhatsnext in a manner that facilitates its transfer to multiple decision contexts.

However, there are ethical risks associated with each potential transfer, in particular the need to ensure the social acceptability, fitness for purpose and appropriate use of readyforwhatsnext in each new decision context [@ethicsCHEMS2024]. For that reason, we have set the goal of readyforwhatsnext meeting six explicit criteria for transparency, reusability and updatability (TRU criteria). We have describe these criteria and their origins in ethical modelling practice elsewhere [@ethicsCHEMS2024], but briefly they are:

@@

U6) Software is systematically retested and deprecated.

### 2024-02-25 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
on to discussion

To enable us to adopt a development approach to readyforwhatsnext that is consistent with meeting all six TRU criteria, we developed a prototype software framework. The software framework is called ready4.

@@

FUR 1) A *template module* for use in CHEM implementations in which modules can: (i) be run as independent models; (ii) safely combined, sharing inputs and outputs with each other without unintended interference; and (iii) be selectively replaced, updated or extended.

FUR 2) A *programming syntax* that promotes simplicity and consistency in how algorithms associated with CHEM modules are described and used in health economic analysis programs.

FUR 3) *Developer tools* that integrate with online software development services to support streamlined workflows for authoring R libaries of CHEM modules that are: (i) created using the template module; (ii) written in a consistent house style; (iii) consistently documented; (iv) licensed for reuse; (v) bundled with citation data specifying authorship, funding source, version number and a Digital Object Identifier (DOI); and (vi) quality assured with publicly available acceptance and unit tests.

@@

FUR 4) *Data management tools* that integrate with online data repository services to streamline the ingesting, labeling and sharing of publicly available and confidential data.

FUR 5) *Reporting tools* that: (i) help author reporting template programs that render documents in HTML, Adobe PDF and Microsoft Word file formats; (ii) retrieve version specific reporting template programs from online code repositories and (iii) supply study specific data (e.g authorship details, input data and analysis results) to reporting template programs to render reports and scientific summaries.
<!-- help generate tables for multiple outputs, retrieve templates, modify templates with authorship and results -->

@@

FUR 7) *Documentation tools* that integrate with open source website development and hosting infrastructure to help: (i) develop a consolidated project documentation website that synthesises content from the individual websites of the project's module libraries, datasets and programs; and (ii) partially automate the updating of website resources such as tutorials and itemised lists of code and data releases.

@@

One library provides the foundation for the ready4 framework. <!-- In modular model implementations, modules need to be able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering].  -->The ready4 foundation library defines a template CHEM module using R’s S4 class system, that supports an object-oriented programming paradigm. Two notable features of object-oriented programming are encapsulation [@8717448] and inheritence [@8717448]. Encapsulation allows a CHEM module to bundle model data and algorithms in a protected environment, which is useful for helping ensure that CHEM modules continue to work as intended when they are combined<!-- [@ready4oop2022] -->. Inheritence allows new CHEM modules to be created that inherit some or all of the properties of a parent module. Inheritence is useful in cases where it is necessary to update CHEM modules that are being combined to account for interaction effects or to selective edit CHEM modulesbeing transfered to different decision contexts<!-- [@ready4oop2022] -->. The ready4 template CHEM module can be used to create other CHEM modules that inherit a common set of properties (meeting *FUR 1*). One of these inherited properties is a novel syntax of 15 core commands that enable CHEM module algorithms to be consistently named (meeting *FUR 2*). The foundation library also contains tools for retrieving web-based information on CHEM modules, datasets and analysis programs authored with ready4 (meeting *FUR 6*) and tools for developing and maintaining a project documentation website (*FUR 7*).

@@

###	Online services 
Framework libraries are designed to be used in conjunction with a number of third-party online services that users of ready4 will need to establish and configure accounts with. The framework's developer tools integrate with online code development, testing and archiving services. The ready4 framework uses the GitHub [@github2007] service's tools for: software development, testing, version control and distribution; hosting of documentation websites; and transparent record keeping of collaborator contributions. The continuous integration [@CI2017] tools provided by GitHub are used to trigger systematic retesting of module libraries with every edit. <!-- Imported -->These tests can be unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). A useful concept for informing software users about the extensiveness of testing procedures is code coverage [@ERICWONG2010188] - reports of the proportion of code that has been explicitly tested. The ready4 framework uses the codecov [@codecov_2022] service to measure and share information about code coverage. <!-- End_Imported -->The open science repository service Zenodo [@Zenodo2013] is used to provide persistent storage, uniquely identified with a a Digital Object Identifier (DOI), of each code release. The framework's data management tools use the Dataverse [@Dataverse2007] data repository service for persistent, uniquely identifiable and versioned storage of model input and output data. For the management of frequently updated datasets that do not require persistent storage, ready4 uses GitHub.

>>>>

To help readyforwhatsnext meet all six TRU criteria, we developed a prototype software framework. The software framework is called ready4 [@ready424].

@@

FUR 1) A reusable *template* for CHEM modules that can: (i) be run as independent models; (ii) safely combined, sharing inputs and outputs with each other without unintended interference; and (iii) be selectively replaced, updated or extended.

FUR 2) A *programming syntax* that promotes simplicity and consistency in how algorithms associated with CHEM modules are: (i) labelled; and (ii) used in health economic analysis programs.

FUR 3) *Module authoring tools* that integrate with online software development services to help author R libraries of CHEM modules that are: (i) created using the template module; (ii) written in a consistent house style; (iii) always at least minimally documented; (iv) licensed for reuse; (v) bundled with authorship, funding source, version number and a Digital Object Identifier (DOI) metadata; and (vi) quality assured with publicly available testing artefacts.

@@

FUR 4) *Data management tools* that integrate with online data repository services to streamline: (i) the ingesting and labeling of confidential and non-confidential data; and  (ii) the sharing of non-confidential data.

FUR 5) *Reporting tools* that: help (i) author reporting templates that render documents in HTML, Adobe PDF and Microsoft Word file formats; (ii) retrieve version specific reporting templates  from online code repositories and (iii) supply study specific data (e.g authorship details, input data and analysis results) to reporting templates to render reports and scientific summaries.
<!-- help generate tables for multiple outputs, retrieve templates, modify templates with authorship and results -->

@@

FUR 7) *Documentation tools* that integrate with open source website development and hosting infrastructure to help: (i) develop a project documentation website that consolidates content from the websites of individual module libraries, datasets and programs; and (ii) partially automate the updating of website resources such as tutorials and itemised lists of code and data releases.

@@

One library provides the foundation for the ready4 framework. <!-- In modular model implementations, modules need to be able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering].  -->The ready4 foundation library defines a template CHEM module using R’s S4 class system, that supports an object-oriented programming paradigm. Two notable features of object-oriented programming are encapsulation [@8717448] and inheritence [@8717448]. Encapsulation allows a CHEM module to bundle model data and algorithms in a protected environment, which is useful for helping ensure that CHEM modules continue to work as intended when combined<!-- [@ready4oop2022] -->. Inheritence allows new CHEM modules to inherit some or all of the properties of a parent module. Inheritence is useful when it is necessary to selectively modify CHEM modules, for example to that to account for interaction effects when CHEMs are combined or to account for the features of a different decision contexts to which they are being transferred<!-- [@ready4oop2022] -->. The ready4 template CHEM module can be used to create other CHEM modules that inherit a common set of properties (meeting *FUR 1*). One of these inherited properties is a novel syntax of 15 core commands that enable CHEM module algorithms to be consistently named (meeting *FUR 2*). The foundation library also contains tools for retrieving web-based information on CHEM modules, datasets and analysis programs authored with ready4 (meeting *FUR 6*) and tools for developing and maintaining a project documentation website (*FUR 7*).

@@

###	Online services 
Framework libraries are designed to be used in conjunction with a number of third-party online services that developers using ready4 need to establish and configure accounts with. The framework's module authoring tools integrate with online code development, testing and archiving services. The ready4 framework uses the GitHub [@github2007] service's tools for: software development, testing, version control and distribution; hosting of documentation websites; and transparent record keeping of collaborator contributions. The continuous integration [@CI2017] tools provided by GitHub are used to trigger systematic retesting of module libraries with every edit. <!-- Imported -->These tests can be unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). A useful concept for informing software users about the extensiveness of testing procedures is code coverage [@ERICWONG2010188] - reports of the proportion of code that has been explicitly tested. The ready4 framework uses the codecov [@codecov_2022] service to measure and share information about code coverage. <!-- End_Imported -->The open science repository service Zenodo [@Zenodo2013] is used to provide persistent storage, uniquely identified with a a Digital Object Identifier (DOI), of each code release. The framework's data management tools use the Dataverse [@Dataverse2007] data repository service for persistent, uniquely identifiable and versioned storage of model input and output data. For the management of frequently updated datasets that do not require persistent storage, ready4 uses GitHub.

### 2024-02-25 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
on to discussion

<!-- Imported_from_Ethics-->Key concerns for health economists when considering whether to reuse a model are generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Health economic models are likely to require at least some modification before they can be validly applied to a new jurisdiction, time period or decision problem. The difficulty of undertaking such a model transfer depends in part on whether the Computational Health Economic Model (CHEM) - the software files that implement the model - can be readily understood, used and edited.

<!-- End_Imported_from_Ethics -->A potential strategy for implementing transferable CHEMs is to adopt a modular and open-source approach to model development. A modular model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. Modules can share inputs and outputs with other modules or can be run as independent models [@barros2023empowering]. Advantages of modular models include feasibility (large projects are broken into smaller tasks, with each component independently developed and tested) and flexibility (making it easier to selectively replace or update specific parts of a model and to scale up or down the level of granularity) [@pan2021modular]. Modular approaches have been used to facilitate the development of complex computational models in disciplines such as biology [@pan2021modular], ecology [@barros2023empowering] and neuroscience [@frazier2022robust]. In health economics the related and enabling concept of reference models has been recommended [@Afzali2013]. Peer reviewed studies describing modular health economic models remain relatively rare, though examples exist in infectious disease [@Trauer2017] and cardiology [@Urach2013].

<!-- Modular CHEMs provide an opportunity for multiple modelling teams to contribute to, test and reuse models.  -->
A number of modular modelling projects have been implemented as open source projects that give others liberal permissions to access and use model source code and data [@pan2021modular; @barros2023empowering; @frazier2022robust]. Many health economists support greater use of open-source CHEMs as a means of improving model transparency, promoting model reuse, making model development more efficient and gaining the confidence of model users [@Pouwels2022]. However, actual implementations of open-source CHEMs are rare [@Emerson2019; @Michalczyk2018; @Feenstra2022], with barriers to adoption including concerns about intellectual property, confidentiality, model misuse and the resources required to maintain open source implementations [@Pouwels2022; @Wu2019]. <!--Benefits --><!-- Adherence to good practice guidance is an essential requirement for healthcare modelling [@Erdemir2020], but  --><!-- Guidelines for implementing open source CHEMs remain scarce, piecemeal and need improving [@Sampson2019].  -->Health economic models can be owned by pharmaceutical companies and consultancies, so commercial considerations can limit the reuse of models and their constituent code and data [@Feenstra2022]. Barriers relating to model generalisability and transferability and the software platforms used to implement CHEMs are also important [@Pouwels2022].

<!-- Transferring a health economic model developed for one jurisdiction for application in another typically involves retaining some features and updating others [@barbieri2010international]. However, models do not always make implementation choices (for example, use of concepts with standardised meanings across jurisdictions)<!-- [@garcia2021cost] -->  <!-- facilitate such selective re-use. -->

<!-- Open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]. The software used to develop CHEMs was identified as a barrier to open-source model implementations by participants in a 2020 survey of health economists [@Pouwels2022]. Other reported barriers relate to difficulties in generalising models, updating models and transferring model data, concerns about the level of public access and lack of interest from decision-makers, and legal, confidentiality and security considerations [@Pouwels2022].  -->

<!-- typically have a number of limitations that at best restrict their usefulness and which in some cases may facilitate harmful misuse. -->

Some of the technical barriers<!-- Give examples --> to the development of transferable CHEMs could be addressed by appropriate software frameworks.<!-- Check last sentence was not used in ethics manuscript  --> A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources (e.g., code and data files), that can be modified to suit specific needs. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python. However, software frameworks can be challenging and time consuming to create and may become excessively complex over time [@edwin2014software]. Software frameworks can also be difficult to learn, often requiring model developers to undergo specialist training [@myllarniemi2018development].

>>>>

<!-- Imported_from_Ethics-->Key concerns for health economists when considering whether to reuse a model are generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Health economic models are likely to require at least some modification before they can be validly applied to a new jurisdiction, time period or decision problem. The difficulty of undertaking such a model transfer depends in part on whether third parties can readily understand, use and edit the software files that implement the model - the Computational Health Economic Model (CHEM).

<!-- End_Imported_from_Ethics -->A potential strategy for implementing transferable CHEMs is to adopt an open-source and modular approach to model development. Open source projects grant third parties liberal permissions to access, use and modify model source code and data.  Many health economists support greater use of open-source CHEMs as a means of improving model transparency, promoting model reuse, making model development more efficient and gaining the confidence of model users [@Pouwels2022]. A number of complex open source computational modelling projects in disciplines such as biology [@pan2021modular], ecology [@barros2023empowering] and neuroscience [@frazier2022robust] have been implemented as modular models. A modular model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. Modules can share inputs and outputs with other modules or can be run as independent models [@barros2023empowering]. Advantages of modular models include feasibility (large projects are broken into smaller tasks, with each component independently developed and tested) and flexibility (making it easier to selectively replace or update specific parts of a model and to scale up or down the level of granularity) [@pan2021modular].

<!-- Modular CHEMs provide an opportunity for multiple modelling teams to contribute to, test and reuse models.  -->
<!-- @Jansen2019 -->
<!-- @thokala2023living LIVING HTA-->
<!-- https://www.ispor.org/publications/journals/value-outcomes-spotlight/vos-archives/issue/view/addressing-assessment-and-access-issues-for-rare-diseases/the-role-of-education-in-shaping-an-open-source-future -->
Actual implementations of open-source CHEMs are relatively rare [@Emerson2019; @Michalczyk2018; @Feenstra2022]. Recent efforts to build this field including an initiative to develop a suite of open source CHEMs addressing different health conditions in the United States [@Jansen2019] and a Canadian platform facilitating access to and validation of multiple CHEMs [@harvard2022developing]. Modular health economic models are also rare, though examples exist in infectious disease [@Trauer2017] and cardiology [@Urach2013]. Barriers to greater use of open-source approaches include health economist concerns about intellectual property, confidentiality, model misuse and the resources required to maintain open source implementations [@Pouwels2022; @Wu2019]. <!--Benefits --><!-- Adherence to good practice guidance is an essential requirement for healthcare modelling [@Erdemir2020], but  --><!-- Guidelines for implementing open source CHEMs remain scarce, piecemeal and need improving [@Sampson2019].  -->When health economic models are owned by pharmaceutical companies and consultancies, commercial considerations may also limit the reuse of models and their constituent code and data [@Feenstra2022]. Health econmists have also identified issues relating to model generalisability and transferability and model implementation software platforms as important barriers to open-source CHEMs [@Pouwels2022]. We are not aware of published literature identifying barriers to modular CHEMs, but potentially these may relate to the limited benefit of undertaking modular approaches in projects that are not also open-source.
<!-- Move to discussion--><!-- In health economics the related and enabling concept of reference models has been recommended [@Afzali2013].  -->
<!-- Transferring a health economic model developed for one jurisdiction for application in another typically involves retaining some features and updating others [@barbieri2010international]. However, models do not always make implementation choices (for example, use of concepts with standardised meanings across jurisdictions)<!-- [@garcia2021cost] -->  <!-- facilitate such selective re-use. -->
<!-- Open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]. The software used to develop CHEMs was identified as a barrier to open-source model implementations by participants in a 2020 survey of health economists [@Pouwels2022]. Other reported barriers relate to difficulties in generalising models, updating models and transferring model data, concerns about the level of public access and lack of interest from decision-makers, and legal, confidentiality and security considerations [@Pouwels2022].  -->
<!-- typically have a number of limitations that at best restrict their usefulness and which in some cases may facilitate harmful misuse. -->

Some of the technical barriers<!-- Give examples --> to the development of transferable CHEMs that use open-source and modular approaches could be addressed by appropriate software frameworks.<!-- Check last sentence was not used in ethics manuscript  --> A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources (e.g., code and data files), that can be modified to suit specific needs. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python. However, software frameworks can be challenging and time consuming to create and may become excessively complex over time [@edwin2014software]. Software frameworks can also be difficult to learn, often requiring model developers to undergo specialist training [@myllarniemi2018development].
